---
title: "Introduction to ggexpress"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggexpress-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ggexpress)
```
When conducting Exploratory Data Analysis it is useful to plot the variables in the data to get an intial sense of the distribution and potential behaviour of the data. If the dataset contains many variables creating separate plots for each one could become tedious. This package automates the plot configuration process and generates basic graphics that summarize the data. 

This package contains 4 functions; two for general purpose exploratory tasks and two that are more specific. 

Creates a basic histogram that indicates the position of the mean and median and displays the standard deviation.

- `gghist`

Creates a scatterplot and calculates the correlation coefficient. 

- `ggscatter`

Creates a Fourier transform plot.

- `fourier_transform`

Converts time series data into 4 subplots displaying the raw data, trend, seasonal and noise components. 

- `ts_plot`

## Data: gapminder
The `gapminder::gapminder` dataset will be used to explore the basic functionality of `gghist` and `ggscatter`. 

```{r}
library(gapminder)
gapminder
```
### Obtain a histogram and basic summary statistics with gghist()
This plot displays the position of the mean and median of life expectancy from the gapminder dataset. In addition, the plot also displays the value of the mean, median and standard deviation. 
```{r gghist_lifeExp, fig.align='center', fig.width=6, fig.height=4}
gghist(data = gapminder, variable = lifeExp)
```
```{r gghist_gdpPercap, fig.align='center', fig.width=6, fig.height=4}
gghist(data = gapminder, variable = gdpPercap)
```

### Obtain a scatter and correlation with ggscatter()

### Visualize a time series and its decomposed components with ts_plot()
```{r ts_plot, fig.align='center', fig.width=6, fig.height=4}
time <- c("1950 Q1", "1950 Q2", "1950 Q3", "1950 Q4", 
          "1951 Q1", "1951 Q2", "1951 Q3", "1951 Q4",
          "1952 Q1", "1952 Q2", "1952 Q3", "1952 Q4")
earnings <- c(0.71, 0.63, 0.82, 0.91, 
              0.51, 1.31, 1.82, 2.01,
              0.63, 0.82, 0.91, 0.71)
ts_data <- tibble::tibble(time, earnings)

ts_plot(ts_data, "earnings", 4)

